\section{lambda}

\begin{lstlisting}[language=Java]
(int a) -> a * 2; <==> a -> a * 2;
(a, b) -> a + b; // Sum of 2 parameters
Multiline lambda use { } and return
(x, y) -> {
	int sum = x + y;
	return sum;
}
\end{lstlisting}

A lambda expression cannot stand alone in Java, it need to be associated to a functional interface.

\begin{lstlisting}[language=Java]
interface MyMath {
    int getDoubleOf(int a);
}
MyMath d = a -> a * 2; // associated to the interface
d.getDoubleOf(4); // is 8
\end{lstlisting}

\section{Collections}

\subsection{sort sort(list, comparator)}
\begin{itemize}
\item{list.sort((a, b)$->$a.length() - b.length())}
\item{list.sort(Comparator.comparing(n$->$n.length()));}
\item{list.sort(Comparator.comparing(String::length));}
\end{itemize}

\subsection{list.\textbf{removeIf}(w$->$w.length()$<$6);}
% //> [Darwin, Galilei, Einstein, Newton]

\subsection{merge merge(key, value, remappingFunction)}

\begin{lstlisting}[language=Java]
Map<String, String> names = new HashMap<>();
names.put("Albert", "Ein?");
// "Albert" exists {Albert=Einstein}
names.merge("Albert", "stein", (old, val) -> old.substring(0, 3) + val);
// not exists{Newname=stein, Albert=Einstein}
names.merge("Newname", "stein", (old, val) -> old.substring(0, 3) + val);
\end{lstlisting}

\section{Class::staticMethod}

Allows to reference methods (and constructors) without executing them

\begin{tabular}{@{}l|l@{}}
    \hline
    {Method Reference}     & {Lambda Form} \\
    \hline
    {StaticMethod.isPrime} & {n$->$StaticMethond.isPrime(n)} \\
    \hline
    {String::toUpperCase}  & {(String w)$->$w.toUpperCase()} \\
    \hline
    {String::compareTo}    & {(String s, String t)$->$s.compareTo(t)}\\
    \hline
    {System.out::println}  & {x$->$System.out.println(x)}\\
    \hline
    {Double::new}          & {n$->$new Double(n)}\\
    \hline
    {String[]::new}        & {(int n)$->$new String[n]}\\
    \hline
\end{tabular}

\section{Streams}

Similar to collections, but
\begin{enumerate}
        \item{They don't store their own data}
        \item{Data comes from elsewhere (collection, file, db, web, ...)}
        \item{immutable (produce new streams)}
        \item{lazy (only computes what is necessary !)}
\end{enumerate}

\begin{lstlisting}[language=Java]
Stream<String> longNames = list
   .filter(n -> n.length() > 8).limit(3);
\end{lstlisting}

\textbf{Creation}
\begin{lstlisting}[language=Java]
Stream<Integer> stream = Stream.of(1, 2, 3, 5);
Stream<String> stream = Stream.of("a", "b");
Stream<String> stream = Stream.of(arr); // array
list.stream(); //list
Stream<Integer> integers = Stream.iterate(0, n -> n + 1);// Infinit stream [0; inf[
\end{lstlisting}

\textbf{Collecting results}

\begin{lstlisting}[language=Java]
// Collect into an array (::new is the constructor reference)
String[] myArray = stream.toArray(String[]::new);
// Collect into a List or Set
List<String> myList = stream.collect(Collectors.toList());
Set<String> mySet = stream.collect(Collectors.toSet());
// Collect into a String
String str = list.collect(Collectors.joining(", "));
\end{lstlisting}

\textbf{map filter reduce, limit, skip}
\begin{lstlisting}[language=Java]
res = stream.map(w -> w.toLowerCase());
res = stream.map(String::toLowerCase);
res = Stream.of(1,2,3,4,5).map(x -> x + 1);
//filter:Retains elements that match
res = stream.filter(n -> n.substring(0, 1).equals("E"));
res = Stream.of(1,2,3,4,5).filter(x -> x < 3);
//Reduce the elements to a single value
String reduced = stream.reduce("", (acc, el) -> acc + "|" + el);
limit limit(maxSize) The n first elements
stream.skip(n) //Discarding the first n elements
stream.distinct() //Remove duplicated elemetns
stream.sorted(); //Must be comparable

//return boolean
boolean res = words.allMatch(n -> n.contains("e"));
//anyMatch: Check if there is a "e" in an element
//noneMatch: Check if there is no "e" in elements
\end{lstlisting}
parallel Returns an equivalent stream that is parallel

findAny faster than findFirst on parallel streams

\subsection{Primitive-Type Streams}
Wrappers (like Stream) are inefficients for unboxing and boxing. Better to use IntStream, DoubleStream, etc.

IntStream stream = IntStream.of(1, 2, 3, 5, 7);
stream = IntStream.of(myArray); // from an array
stream = IntStream.range(5, 80); // range from 5 to 80

Random gen = new Random();
IntStream rand = gen(1, 9); // stream of randoms
Use mapToX (mapToObj, mapToDouble, etc.) if the function yields Object, double, etc. values.

Grouping Results

Collectors.groupingBy

// Groupe by length
Map<Integer, List<String>> groups = stream
	.collect(Collectors.groupingBy(w -> w.length()));
//> 4=[Bohr], 5=[Tesla], 6=[Darwin, Newton], ...
Collectors.toSet

// Same as before but with Set
... Collectors.groupingBy(
	w -> w.substring(0, 1), Collectors.toSet()) ...
Collectors.counting Count the number of values in a group

Collectors.summing\_\_ summingInt, summingLong, summingDouble to sum group values

Collectors.averaging\_\_ averagingInt, averagingLong, ...

// Average length of each element of a group
Collectors.averagingInt(String::length)
PS: Don't forget Optional (like Map<T, Optional<T>>) with some Collection methods (like Collectors.maxBy).

Parallel Streams

Creation

Stream<String> parStream = list.parallelStream();
Stream<String> parStream = Stream.of(myArray).parallel();
unordered Can speed up the limit or distinct

stream.parallelStream().unordered().distinct();
PS: Work with the streams library. Eg. use filter(x -> x.length() < 9) instead of a forEach with an if.

Optional

In Java, it is common to use null to denote absence of result. Problems when no checks: NullPointerException.

// Optional<String> contains a string or nothing
Optional<String> res = stream
   .filter(w -> w.length() > 10)
   .findFirst();

// length of the value or "" if nothing
int length = res.orElse("").length();

// run the lambda if there is a value
res.ifPresent(v -> results.add(v));
Return an Optional

Optional<Double> squareRoot(double x) {
   if (x >= 0) { return Optional.of(Math.sqrt(x)); }
   else { return Optional.empty(); }
}
Note on inferance limitations

interface Pair<A, B> {
    A first();
    B second();
}
A steam of type Stream<Pair<String, Long>> :

stream.sorted(Comparator.comparing(Pair::first)) // ok
stream.sorted(Comparator.comparing(Pair::first).thenComparing(Pair::second)) // dont work
Java cannot infer type for the .comparing(Pair::first) part and fallback to Object, on which Pair::first cannot be applied.

The required type for the whole expression cannot be propagated through the method call (.thenComparing) and used to infer type of the first part.

Type must be given explicitly.

stream.sorted(
    Comparator.<Pair<String, Long>, String>comparing(Pair::first)
    .thenComparing(Pair::second)
)
